FCST

### 2021-03-11 analysis/chapters/forecast_20181122.Rmd
Graficos de resumen de RMSE y BIAS

```{r message=FALSE, warning=FALSE}
files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
files <- files[str_detect(files, "_01_")]

sondeos <- purrr::map(files, function(f){

  fread(f) %>% 
    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
    .[, launch_time := as_datetime(launch_time)] %>% 
    .[, fcst_exp := basename(dirname(dirname(f)))]

@@

sondeos %>% 
  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
  .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
  .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
           BIAS = mean(value - fcst_value, na.rm = TRUE)), 
    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
  ggplot(aes(lev/1000, value.est)) +
  geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()
```

```{r rmse-sondeos-media}
extremos <- miembros %>% 
  .[, .(max_rmse = max(RMSE),
        min_rmse = min(RMSE)), by = .(exp, fcst_exp, lev, variable)]

@@

```{r rmse-sondeos-param}
pbl <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, lev, variable)]

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(pbl = "Mean",
           RMSE = value.est)] %>% 
  .[, value.est := NULL] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

cum <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, lev, variable)]

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(cumulus = "Mean",
           RMSE = value.est)] %>% 
  .[, value.est := NULL] %>% 
  rbind(., cum) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

@@

```{r bias-sondeos-media}
extremos <- miembros %>% 
  .[, .(max_bias = max(BIAS),
        min_bias = min(BIAS)), by = .(exp, fcst_exp, lev, variable)]

@@

```{r bias-sondeos-param}
pbl <- miembros[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, lev, variable)]

media %>% 
  .[estadistico == "BIAS"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
  .[, ":="(pbl = "Mean",
           BIAS = value.est)] %>% 
  .[, value.est := NULL] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(lev/1000, BIAS)) +
  geom_line(aes(y = BIAS, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

cum <- miembros[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, lev, variable)]

media %>% 
  .[estadistico == "BIAS"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
  .[, ":="(cumulus = "Mean",
           BIAS = value.est)] %>% 
  .[, value.est := NULL] %>% 
  rbind(., cum) %>% 
  ggplot(aes(lev/1000, BIAS)) +
  geom_line(aes(y = BIAS, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

@@

media <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean2.rds")),
               read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean2.rds")))

miembros <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds")),
                  read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds")))

@@

```{r}
cum <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, var, date)]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
  .[, ":="(cumulus = "Mean")] %>% 
  rbind(., cum) %>% 
  ggplot(aes(date, BIAS)) +
  geom_line(aes(color = exp, linetype = cumulus)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_date() +
  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

cum <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, var, date)]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
  .[, ":="(cumulus = "Mean")] %>% 
  rbind(., cum) %>% 
  ggplot(aes(date, RMSE)) +
  geom_line(aes(color = exp, linetype = cumulus)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  # geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()
```
```{r}
pbl <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, var, date)]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
  .[, ":="(pbl = "Mean")] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(date, BIAS)) +
  geom_line(aes(color = exp, linetype = pbl)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_date() +
  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

pbl <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, var, date)]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
  .[, ":="(pbl = "Mean")] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(date, RMSE)) +
  geom_line(aes(color = exp, linetype = pbl)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  # geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()
```

>>>>

```{r message=FALSE, warning=FALSE}
# files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
# files <- files[str_detect(files, "_01_")]
# 
# sondeos <- purrr::map(files, function(f){
#   
#   fread(f) %>% 
#     .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
#     .[, launch_time := as_datetime(launch_time)] %>% 
#     .[, fcst_exp := basename(dirname(dirname(f)))]
#   
# }) %>% 
#   rbindlist()

@@

media %>% 
  ggplot(aes(lev/1000, value.est)) +
  geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()
```

```{r rmse-sondeos-ensamble}
extremos <- miembros %>% 
  .[, .(max_rmse = max(RMSE),
        min_rmse = min(RMSE)), by = .(exp, fcst_exp, lev, variable)]

@@

```{r rmse-sondeos-param}
pbl <- miembros[, .(RMSE = mean(RMSE),
                    tipo = "pbl"), 
                by = .(exp, fcst_exp, pbl, lev, variable)] %>% 
  setnames("pbl", "param")

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           RMSE = value.est, 
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

cum <- miembros[, .(RMSE = mean(RMSE),
                    tipo = "cumulus"), 
                by = .(exp, fcst_exp, cumulus, lev, variable)] %>% 
  setnames("cumulus", "param")

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           RMSE = value.est,
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., cum) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

rbind(pbl, cum) %>% 
  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
    by = .(variable, exp, param, fcst_exp)] %>% 
  ggplot(aes(RMSE_mean)) +
  geom_histogram(aes(fill = exp), 
                 position = "identity", 
                 alpha = 0.5) +
   scale_fill_manual(values = colores_exp, 
                     labels = c(E4 = "CONV", E5 = "AUT",
                                E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ variable, scales = "free_x")

rbind(pbl, cum) %>%
  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
    by = .(variable, exp, param, fcst_exp, tipo)] %>% 
  ggplot(aes(RMSE_mean, exp)) +
  geom_point(aes(color = exp, shape = param),
             size = 3) +
  scale_color_manual(guide = NULL, values = colores_exp, 
                      labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ variable, scales = "free_x") +
  labs(y = NULL, x = "mean RMSE", shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        panel.background = element_rect(fill = "grey98",
                                        color = "grey98"))

@@

```{r bias-sondeos-ensamble}
extremos <- miembros %>% 
  .[, .(max_bias = max(BIAS),
        min_bias = min(BIAS)), by = .(exp, fcst_exp, lev, variable)]

@@

```{r bias-sondeos-param}
pbl <- miembros[, .(BIAS = mean(BIAS),
                    tipo = "pbl"), 
                by = .(exp, fcst_exp, pbl, lev, variable)] %>% 
  setnames("pbl", "param")

media %>% 
  .[estadistico == "BIAS"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           BIAS = value.est,
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(lev/1000, BIAS)) +
  geom_line(aes(y = BIAS, color = exp, linetype = param, group = interaction(exp, param))) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

cum <- miembros[, .(BIAS = mean(BIAS),
                    tipo = "cumulus"), 
                by = .(exp, fcst_exp, cumulus, lev, variable)] %>% 
  setnames("cumulus", "param")

media %>% 
  .[estadistico == "BIAS"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           BIAS = value.est,
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., cum) %>% 
  ggplot(aes(lev/1000, BIAS)) +
  geom_line(aes(y = BIAS, color = exp, linetype = param, group = interaction(exp, param))) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

rbind(pbl, cum) %>%
  .[, .(BIAS_mean = mean(BIAS, na.rm = TRUE)), 
    by = .(variable, exp, param, fcst_exp, tipo)] %>% 
  ggplot(aes(BIAS_mean, exp)) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_point(aes(color = exp, shape = param),
             size = 3) +
  scale_color_manual(guide = NULL, values = colores_exp, 
                      labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ variable, scales = "free_x") +
  labs(y = NULL, x = "mean BIAS", shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        panel.background = element_rect(fill = "grey98",
                                        color = "grey98"))

@@

media <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean2.rds")),
               read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean2.rds"))) %>% 
  .[, ":="(RMSE = fifelse(var == "q", RMSE*1000, RMSE),
           BIAS = fifelse(var == "q", BIAS*1000, BIAS))]

miembros <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds")),
                  read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds"))) %>% 
    .[, ":="(RMSE = fifelse(var == "q", RMSE*1000, RMSE),
           BIAS = fifelse(var == "q", BIAS*1000, BIAS))]

@@

```{r}
cum <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, var, date)] %>% 
   setnames("cumulus", "param") %>% 
  .[, tipo := "cumulus"]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
  .[, ":="(param = "Mean",
           tipo = "Mean")] %>% 
  rbind(., cum) %>% 
  ggplot(aes(date, BIAS)) +
  geom_line(aes(color = exp, linetype = param)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_date() +
  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

pbl <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, var, date)] %>% 
   setnames("pbl", "param") %>% 
  .[, tipo := "pbl"]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
  .[, ":="(param = "Mean",
           tipo = "Mean")] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(date, BIAS)) +
  geom_line(aes(color = exp, linetype = param)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  # geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_date() +
  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

rbind(pbl, cum) %>%
  .[, .(BIAS_mean = mean(BIAS, na.rm = TRUE)), 
    by = .(var, exp, param, fcst_exp, tipo)] %>% 
  ggplot(aes(BIAS_mean, exp)) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_point(aes(color = exp, shape = param),
             size = 3) +
  scale_color_manual(guide = NULL, values = colores_exp, 
                      labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ var, scales = "free_x") +
  labs(y = NULL, x = "mean BIAS", shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        panel.background = element_rect(fill = "grey98",
                                        color = "grey98"))
```

```{r}
cum <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, var, date)] %>% 
   setnames("cumulus", "param") %>% 
  .[, tipo := "cumulus"]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
  .[, ":="(param = "Mean",
           tipo = "Mean")] %>% 
  rbind(., cum) %>% 
  ggplot(aes(date, RMSE)) +
  geom_line(aes(color = exp, linetype = param)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

pbl <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, var, date)] %>% 
   setnames("pbl", "param") %>% 
  .[, tipo := "pbl"]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
  .[, ":="(param = "Mean",
           tipo = "Mean")] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(date, RMSE)) +
  geom_line(aes(color = exp, linetype = param)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  # geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

rbind(pbl, cum) %>%
  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
    by = .(var, exp, param, fcst_exp, tipo)] %>% 
  ggplot(aes(RMSE_mean, exp)) +
  geom_point(aes(color = exp, shape = param),
             size = 3) +
  scale_color_manual(guide = NULL, values = colores_exp, 
                      labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ var, scales = "free_x") +
  labs(y = NULL, x = "mean RMSE", shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        panel.background = element_rect(fill = "grey98",
                                        color = "grey98"))
```
